---
title: "REVERSAL TASK PROCESSING"
author: "Michelle.VanTieghem"
date: "January 27, 2016"
output: slidy_presentation
---

```{r Read in data files, echo = FALSE}
##########################################s
# Making a list of file names and checking number
#########################################

# setwd ("~/Documents/Documents/Columbia/LabDATA/Reversal_task/Reversal/analysis/")
file_dir = "C:/Users/maria/Desktop/SLCN/reversal4choicegame/logs"
output_dir = "C:/Users/maria/Desktop/SLCN/reversal4choicegame/analysis"
subject.files = list.files(path = file_dir,
                           pattern = "logfile")
length(subject.files) #64
subject.list = substr(subject.files, 1, 80)
subject.list 

#########################################
# Load all files for each subject and loop 
#########################################

# create empty listto store each subjects data frame 
data.list = list()  
# data.file

for(i in (subject.list)){

#  i = subject.list[1]
  # create file path to loop through all subjects 
  filepath = file.path(file_dir, paste(i, sep = ""))
  # assign data frame for each subject 
  data.file = assign(i, read.table(filepath, stringsAsFactors=T, header=T) # PROBLEM: CURRENT TIME TAKES UP TWO ROWS!!, row.names = NULL))
  data.file = data.frame(data.file)
  # add data for each subject to list of data frames 
  data.list[[i]] = data.file

  }

# check names and then split into chunks

length(data.list) #64
names(data.list)

#############################################
## make empty dataframe to store sub data
#############################################

compiled.data =  data.frame("subject" = numeric(0),"Age" = numeric(0), "gender" = numeric(0), "Ave.choiceonseTime" = numeric(0),  "reversal.complete" = numeric(0),"trials.learn.aq" = numeric(0), "trials.learn.rev"= numeric(0), "persev.choice.rev"= numeric(0) , "error.choice.rev" = numeric(0))

#############################################
## start the data loop  for all subs
#######################s######################
for (i in 1:length(data.list)){
  # i = 3
  data.set = data.list[[i]]
  subject = subject.list[i]
  Age = as.numeric(data.set$Age[1])
  gender = as.character(data.set$Gender[1])
  data.length = nrow(data.set)
 
#*#*#* need to add one here because Python starts at 0
  TotalTrials = data.set$TrialNumber[data.length]+1

##################################################################
## PROCESSING FOR AQ PHASE
##################################################################
  trials.learn.aq = 0
  for (i in (1:TotalTrials)){
    if(data.set$Reversal[i] == 0){
      trials.learn.aq = trials.learn.aq+1
    }
  }
  trials.learn.aq
  ## visits boxes - on average ... 
  target.aq = data.set$Target[1]
  persev.prior.choice.aq  = 0 
  error.choice.aq = 0
  rewarded.choice.aq = 0
  Visits0.aq = 0
  Visits1.aq = 0
  Visits2.aq = 0
  Visits3.aq = 0
  
  ###################################################
  ### while loop for AQ phase
  #######################################################
  k =1
  while (k >= 1 & k<= trials.learn.aq){
    print (k)
    
    #### VISITS 
    # for each trial, add number of visits per box
    Visits0.aq = Visits0.aq + data.set$Visits0[k]
    Visits1.aq = Visits1.aq + data.set$Visits1[k]
    Visits2.aq = Visits2.aq + data.set$Visits2[k]
    Visits3.aq = Visits3.aq + data.set$Visits3[k]
     
  
    ### CHOICES 
    if(data.set$Choice[k] == target.aq){
      rewarded.choice.aq = rewarded.choice.aq + 1
     } 
    if(data.set$Choice[k] != target.aq) {
      error.choice.aq = error.choice.aq +1 
      }
    if(k > 1){
      j = k-1
      if(data.set$Choice[k] == data.set$Choice[j]){
       persev.prior.choice.aq = persev.prior.choice.aq + 1
      }}
    k = k + 1
    } 
  
  error.choice.aq
  rewarded.choice.aq
  persev.prior.choice.aq
  
  Visits0.aq
  Visits1.aq 
  Visits2.aq
  Visits3.aq 
  #######################################
  ### calculating visit parameters
  ####################################
 total.reward.visits.aq = 0
 total.incorrect.visits.aq = 0
 total.visits.aq = 0
  
   if (target.aq == 0){
     total.reward.visits.aq = Visits0.aq
       total.incorrect.visits.aq = Visits1.aq + Visits2.aq + Visits3.aq
   } else if(target.aq == 1){
     total.reward.visits.aq = Visits1.aq
       total.incorrect.visits.aq = Visits0.aq + Visits2.aq + Visits3.aq
   }else if(target.aq == 2){
     total.reward.visits.aq = Visits2.aq
       total.incorrect.visits.aq = Visits1.aq + Visits0.aq + Visits3.aq
   }else if(target.aq == 3){
     total.reward.visits.aq = Visits3.aq
       total.incorrect.visits.aq = Visits1.aq + Visits2.aq + Visits0.aq
   }
  
                             
  visits.total.aq = sum(Visits0.aq, Visits1.aq, Visits2.aq, Visits3.aq)
  visits.total.aq
  check = ifelse((total.incorrect.visits.aq + total.reward.visits.aq) == visits.total.aq, "CHECK!", "ERROR!")
  check
  
  
  
  ######################################################
  ## find number of trials to learn REV 
  ######################################################
  trials.learn.rev = 0
  for (i in (1:TotalTrials)){
    print(data.set$Reversal[i])
    if(data.set$Reversal[i]== 1){
      trials.learn.rev = trials.learn.rev+1
    }
  }
  
  trials.learn.rev
  
  ##########################################################
  ## get rid of subjects who don't finish task and mark NAs
  ######################################################
  
  ## didn't finish REV 
  if (trials.learn.rev > 0  &  data.set$RunningTotal[TotalTrials] >= 4){
    learned.rev = "YES"
  } else {learned.rev = "NO"}
  
  learned.rev
  
  ## didn't finish AQ 
  if(trials.learn.rev == 0 | trials.learn.aq  > 30){
    learned.aq = "NO"
    switch.trial = "NA"
    target.rev = "NA"
    persev.choice.rev  = "NA"
    error.choice.rev = "NA"
    rewarded.choice.rev = "NA"
    
  ##########################################################
  ## keep  subjects who did finish task and continue
  ######################################################
  } else if(trials.learn.rev > 0){
    
    ### setting up all variable names 
    learned.aq = "YES"
    switch.trial = trials.learn.aq +1
    Visits0.rev = 0
    Visits1.rev = 0
    Visits2.rev = 0
    Visits3.rev = 0
    target.rev = data.set$Target[switch.trial] 
    novel.rev = 3 # always 3! 
    novel.choice.rev = 0
    persev.choice.rev  = 0 
    error.choice.rev = 0
    rewarded.choice.rev = 0
    irrelevant.choice.rev = 0
  
    ######################################
    ### starting while loop for reversal
    ##########################################
    p = switch.trial
    while (p >= switch.trial & p<= TotalTrials){
      print (p)
       
      # for each trial, add number of visits per box
      Visits0.rev = Visits0.rev + data.set$Visits0[p]
      Visits1.rev = Visits1.rev + data.set$Visits1[p]
      Visits2.rev = Visits2.rev + data.set$Visits2[p]
      Visits3.rev = Visits3.rev + data.set$Visits3[p]
       
      ### Choices 
      if(data.set$Choice[p] == target.rev){
        rewarded.choice.rev = rewarded.choice.rev + 1
      } 
      if(data.set$Choice[p] != target.rev) {
        error.choice.rev = error.choice.rev +1 
        }
      if(data.set$Choice[p] == target.aq){
         persev.choice.rev = persev.choice.rev + 1
        } 
      if (data.set$Choice[p] == 3){
        novel.choice.rev = novel.choice.rev + 1
      }
      if (data.set$Choice[p] != target.aq & data.set$Choice[p] != target.rev & data.set$Choice[p] != 3){
        irrelevant.choice.rev = irrelevant.choice.rev + 1
      }
      p =p + 1
      } 
    }
  #######################################
  ### calculating visit parameters
  ####################################
     total.reward.visits.rev = 0
     total.persev.visits.rev = 0
     total.irrelevant.visits.rev = 0
     total.novel.visits.rev =0
     total.visits.rev = 0
    
   if (target.rev == 0){
     total.reward.visits.rev = Visits0.rev
   } else if(target.rev == 1){
     total.reward.visits.rev = Visits1.rev
   }else if(target.rev == 2){
     total.reward.visits.rev = Visits2.rev
   }else if(target.rev == 3){
     total.reward.visits.rev = Visits3.rev
   }
  
   if (target.aq == 0){
     total.persev.visits.rev = Visits0.rev
   } else if(target.aq == 1){
     total.persev.visits.rev = Visits1.rev
   }else if(target.aq == 2){
     total.persev.visits.rev = Visits2.rev
   }else if(target.aq == 3){
     total.persev.visits.rev = Visits3.rev
   }
    
   
   if (novel.rev == 0){
     total.novel.visits.rev = Visits0.rev
   } else if(novel.rev == 1){
     total.novel.visits.rev = Visits1.rev
   }else if(novel.rev == 2){
     total.novel.visits.rev = Visits2.rev
   }else if(novel.rev == 3){
     total.novel.visits.rev =Visits3.rev
   }
  
     # calculate irrelevant box number ... s
     box.list = cbind(1,2,3,4)
     for (i in box.list){
       current.i = box.list[i]-1
       if (current.i != (target.rev) & current.i != (target.aq) & current.i != (novel.rev)){
         irrelevant.rev = current.i
       }
     }
  
   if (irrelevant.rev == 0){
     total.irrelevant.visits.rev = Visits0.rev
   } else if(irrelevant.rev == 1){
     total.irrelevant.visits.rev = Visits1.rev
   } else if(irrelevant.rev == 2){
     total.irrelevant.visits.rev = Visits2.rev
   } else if(irrelevant.rev == 3){
     total.irrelevant.visits.rev = Visits3.rev
   }
  
  visits.incorrect.rev = sum(total.irrelevant.visits.rev, total.persev.visits.rev, total.novel.visits.rev)
  visits.incorrect.rev 
  total.reward.visits.rev 
  visits.total.rev = sum(Visits0.rev, Visits1.rev, Visits2.rev, Visits3.rev)
  visits.total.rev 
  check = ifelse((visits.incorrect.rev + total.reward.visits.rev) == visits.total.rev, "CHECK!", "ERROR!")
  check
  
  
  ######################################################
  ## calculating RT parameters
  #####################################################
  
  rt.aq = mean(ifelse(data.set$Reversal == 0, data.set$ResponseTime, NA), na.rm = TRUE)
  rt.reward.aq = mean(ifelse((data.set$Reversal == 0 & data.set$Choice == target.aq), data.set$ResponseTime, NA), na.rm = TRUE)
  rt.error.aq = mean(ifelse((data.set$Reversal == 0 & data.set$Choice != target.aq), data.set$ResponseTime, NA), na.rm = TRUE)
  rt.rev = mean(ifelse(data.set$Reversal == 1, data.set$ResponseTime, NA), na.rm = TRUE)
  rt.reward.rev = mean(ifelse((data.set$Reversal == 1 & data.set$Choice == target.aq), data.set$ResponseTime, NA), na.rm = TRUE)
  rt.error.rev = mean(ifelse((data.set$Reversal == 1 & data.set$Choice != target.aq), data.set$ResponseTime, NA), na.rm = TRUE)
  rt.average = mean(data.set$ResponseTime, na.rm = TRUE)
  
  ############################################################
  ## combine all data and add to giant dataframe
  ############################################################
  sub.row = cbind(subject, Age, gender, TotalTrials, 
  target.aq, target.rev, novel.rev, irrelevant.rev,
  learned.aq,  trials.learn.aq, rewarded.choice.aq,error.choice.aq,  persev.prior.choice.aq, 
  learned.rev, trials.learn.rev, rewarded.choice.rev, error.choice.rev, persev.choice.rev, novel.choice.rev,
  Visits0.aq, Visits1.aq, Visits2.aq, Visits3.aq, 
  Visits0.rev, Visits1.rev, Visits2.rev, Visits3.rev, 
  total.reward.visits.aq, total.error.visits.aq, 
  total.reward.visits.rev, total.novel.visits.rev, total.persev.visits.rev, total.irrelevant.visits.rev,
  rt.aq, rt.reward.aq, rt.error.aq, rt.rev, rt.reward.rev, rt.error.rev, rt.average)
  sub.row  
  
  compiled.data = data.frame(structure(rbind(compiled.data, sub.row)))
  #compiled.test = data.frame(rbind(compiled.data, sub.row))
}
  
head(compiled.data)
nrow(compiled.data)
file.name = paste0(output_dir, "N64.Reversal.compiled.data.csv")
write.csv(compiled.data, file.name)
output_dir
```

############################################################
## analysis with age!!
############################################################

```{r, echo = FALSE}
head(compiled.data)
nrow(compiled.data)
# how many subjects completed the task.
summary(as.factor(compiled.data$learned.aq))
summary(as.factor(compiled.data$learned.rev)) #

# average age and histogram check
compiled.data$Age.real = as.numeric(as.character(compiled.data$Age))
hist(compiled.data$Age.real)
summary((compiled.data$Age.real)) # min = 4, mean is 13
```

### Filter data for completed data 
```{r, echo = FALSE}


#### ONLY LOOKING AT COMLETED
complete.rev = subset(compiled.data, learned.rev == "YES")
nrow(complete.rev)
summary(complete.rev$Age.real) 
hist(complete.rev$Age.real)

```


### number of trials to criterion for AQ  & REV 
```{r, echo = FALSE }

complete.rev$TotalTrials = as.numeric(complete.rev$TotalTrials)
hist(complete.rev$TotalTrials)
summary(complete.rev$TotalTrials)

complete.rev$trials.learn.aq
hist(as.numeric(complete.rev$trials.learn.aq))
summary(complete.rev$trials.learn.aq)

complete.rev$trials.learn.rev = as.numeric(complete.rev$trials.learn.rev)
hist(complete.rev$trials.learn.rev)
summary(complete.rev$trials.learn.rev)


complete.rev$rewarded.choice.aq = as.numeric(complete.rev$rewarded.choice.aq)
hist(complete.rev$rewarded.choice.aq)
summary(complete.rev$rewarded.choice.aq)

complete.rev$error.choice.aq = as.numeric(complete.rev$error.choice.aq)
hist(complete.rev$error.choice.aq)
summary(complete.rev$error.choice.aq)


complete.rev$rewarded.choice.rev = as.numeric(complete.rev$rewarded.choice.rev)
hist(complete.rev$rewarded.choice.rev)
summary(complete.rev$rewarded.choice.rev)

complete.rev$error.choice.rev = as.numeric(complete.rev$error.choice.rev)
hist(complete.rev$error.choice.rev)
summary(complete.rev$error.choice.rev)



################################
 # REVERSAL
################################s

compiled.data$trials.learn.rev = as.numeric(as.character(compiled.data$trials.learn.rev))
hist(compiled.data$trials.learn.rev) # minimum should be 5 
summary(compiled.data$trials.learn.rev) # min shouldn't be zero?! there aren't any NA so this should be rewarded.

compiled.data$rewarded.choice.rev = as.numeric(as.character(compiled.data$rewarded.choice.rev))
hist(compiled.data$rewarded.choice.rev) # minimum should be 5 
summary(compiled.data$rewarded.choice.rev) # why is the max so large?!  

compiled.data$error.choice.rev = as.numeric(as.character(compiled.data$error.choice.rev))
summary((compiled.data$error.choice.rev)) # minimum rewarded =5 , GOOD! 
# and here there are NAs which means that worked.
hist(compiled.data$error.choice.rev)


compiled.data.incomplete.aq = subset(compiled.data, learned.aq == "NO")
compiled.data.complete.aq = subset(compiled.data, learned.aq == "YES")
compiled.data.incomplete.rev = subset(compiled.data, learned.rev == "NO")
hist(compiled.data.incomplete.rev$Age.real)
hist(compiled.data.incomplete.aq$Age.real) # few kids older never learned rev.
summary(compiled.data.incomplete.aq$Age.real)
nrow(complete.rev)
plot(complete.rev$Age.real, complete.rev$trials.learn.aq)
plot(complete.rev$Age.real, complete.rev$trials.learn.rev)
plot(complete.rev$Age.real, complete.rev$persev.prior.choice.aq)
plot(complete.rev$Age.real, complete.rev$persev.choice.rev)

complete.rev$age.bins = ifelse(complete.rev$Age.real >= 30, "30-40", ifelse(complete.rev$Age.real >=25, "25-29", ifelse(complete.rev$Age.real >= 20, "20-24", ifelse(complete.rev$Age.real >= 15, "15-19", ifelse(complete.rev$Age.real >=10, "10-14",ifelse(complete.rev$Age.real >= 6, "6-9", NA) )))))
summary(as.factor(complete.rev$age.bins))

